Flaring star-planet interactions in star-planet systems in Kepler and TESS

### 2023-05-12 src/tex/ms.tex
[showyourwork] overleaf sync

We collect the planetary and stellar masses from the literature~(Table~\ref{tab:tidal}). For missing planetary masses, the radii were known, so used the mass-radius relations implemented in \texttt{forecaster}~\citep{chen2017probabilistic}. GJ 3323, GJ 674 and GJ 3082 were also missing stellar mass estimates, which we obtained from the \citep{mann2015how, mann2016erratum} relations between absolute Ks magnitude~\citep{skrutskie2006two} and stellar mass for M dwarfs using the distances from \citet{bailer-jones2021estimating}. We list the masses in Table~\ref{tab:tidal}. For Kepler-42 b, we used  $M_p=0.1-2.06M_\oplus$ that covers pure rock to pure iron compositions in \citet{muirhead2012characterizing}, and adopted the logarithmic mean for our estimate. Kepler-1558 c has a similar radius to Kepler-42 b, i.e., most likely a bare core planet, so we proceeded analogously to estimate its mass.

@@

\paragraph{KOI-12}, which lacks signal with $P_{orb}$ despite high expected power of magnetic interaction, shows elevated expected and measured tidal interaction. This could be interpreted as KOI-12 b orbiting outside the Alfv\'en radius, unable to interact magnetically, but tidally. We note that KOI may be evolving off the main sequence, so that the tidal interaction estimates might not be accurate for this star.

\paragraph{TAP 26,} which shows no signs of magnetic SPI, does not show any sign of tidal interaction despite both being high according to the scaling laws. This could again be due to the inclination of TAP 26's orbit, but only if the flares caused by the tidal bulge are not close to the equator. Otherwise, TAP 26's tidally induced flares should be more modulated than the magnetically induced ones.

\paragraph{Kepler-42 and TOI-540}. Kepler-42 is neither expected to interact magnetically, nor did it show any deviation from random flare timing in phase with $P_{orb}$. However, since it is the system with the shortest planetary orbit in our sample, it is both expected to and does show marginal signal of tidal interaction. Kepler-42 is very similar to TOI-540, which also is a small M dwarf with a likely terrestrial planet in a very short orbit. A key difference is that Kepler-42 is rotating at about 70 days, and therefore expected to have a relatively weak magnetic field, whereas TOI-540 is a young star with a rotation period $<1\,$d. TOI-540 is expected to have lower tidal interaction strength than Kepler-42 due to TOI-540 b's wider orbit, and also shows a weaker tidal flaring SPI signal in all three scenarios. K2-25 and AU Mic are similar, but differ from TOI-540 -- both show tentative signs of magnetic SPI, but are consistent within $1\sigma$ with no observable tidal interaction.

>>>>

We collect the planetary and stellar masses from the literature~(Table~\ref{tab:tidal}). For missing planetary masses, the radii were known, so used the mass-radius relations implemented in \texttt{forecaster}~\citep{chen2017probabilistic}. GJ 3323, GJ 674 and GJ 3082 were also missing stellar mass estimates, which we obtained from the \citep{mann2015how, mann2016erratum} relations between absolute Ks magnitude~\citep{skrutskie2006two} and stellar mass for M dwarfs using the distances from \citet{bailer-jones2021estimating}. We list the masses in Table~\ref{tab:tidal}. For Kepler-42 b, we used  $M_p=0.1-2.06M_\oplus$ that covers pure rock to pure iron compositions in \citet{muirhead2012characterizing}, and adopted the logarithmic mean for our estimate. Kepler-1558 c has a similar radius to Kepler-42 b, i.e., most likely a bare core planet, so we proceeded analogously to estimate its mass. We also adopt the upper limit of $5M_J$ for HIP 67522, but use the lower error from \citet{chen2017probabilistic}.

@@

\paragraph{KOI-12,} which lacks signal with $P_{orb}$ despite high expected power of magnetic interaction, shows elevated expected and measured tidal interaction. This could be interpreted as KOI-12 b orbiting outside the Alfv\'en radius, unable to interact magnetically, but tidally. We note that KOI-12 may be evolving off the main sequence, so that the tidal interaction estimates might not be accurate for this star.

\paragraph{TAP 26,} which shows no signs of magnetic SPI, does not show any sign of tidal interaction, either, despite both being high according to the scaling laws. This could again be due to the inclination of TAP 26's orbit, but only if the flares caused by the tidal bulge are not close to the equator. Otherwise, TAP 26's tidally induced flares should be more modulated than the magnetically induced ones.

\paragraph{Kepler-42 and TOI-540} Kepler-42 is neither expected to interact magnetically, nor did it show any deviation from random flare timing in phase with $P_{orb}$. However, since it is the system with the shortest planetary orbit in our sample, it is both expected to and does show marginal signal of tidal interaction. Kepler-42 is very similar to TOI-540, which also is a small M dwarf with a likely terrestrial planet in a very short orbit. A key difference is that Kepler-42 is rotating at about 70 days, and therefore expected to have a relatively weak magnetic field, whereas TOI-540 is a young star with a rotation period $<1\,$d. TOI-540 is expected to have lower tidal interaction strength than Kepler-42 due to TOI-540 b's wider orbit, and also shows a weaker tidal flaring SPI signal in all three scenarios. K2-25 and AU Mic are similar, but differ from TOI-540 -- both show tentative signs of magnetic SPI, but are consistent within $1\sigma$ with no observable tidal interaction.
